---
title: "Homework 8 - Multiple Linear Regression"
author: "Austin Chan"
date: "April 21, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
---

##8.2

###a).

$Weight = -1.93*Parity + 120.07$

###b).

First born babies have a predicted birth weight 1.93 ounces lower than babies who are not first born.

###c).

Since the p-value for the slope is >0.1 at the confidence level of 0.05, there is not a statistically significant relationship between average birth weight and parity.

##8.4

###a).

$AbsentDays = -9.11*Ethnicity + 3.10*Sex + 2.15*Learner + 18.93$

###b).

Ethnicity - All else being constant, the predicted number of absent days for not aboriginal students is 9.11 days lower than aboriginal students.

Sex - All else being constant, the predicted number of absent days for males is 3.1 days higher than females.

Learner Status - All else being constant, the predicted number of absent days for slow learners is 2.15 days higher than average learners.

###c).

```{r}
predicted = -9.11*0 + 3.10*1 + 2.15*1 + 18.93
observed = 2

predicted - observed
```

###d).

```{r}
Rsquared = 1 - (240.57/264.17)
RsquaredAdj = 1 - ((240.57/(146 - 3 - 1))/(264.17/(146-1)))

Rsquared
RsquaredAdj
```

##8.8

Learner status should be removed because it increases the adjusted Rsquared when it is removed.

##8.16

###a).

The lower the temperature, the higher the likelihood that there were damaged O-rings.

###b).

Intercept - The estimated number of O-ring failures when the temperature is zero is 11.66.

Temperature - As temperature increases by one degree, the estimated number of O-ring failures decreases by 0.2162.

###c).

$log(phat/1-phat)  = -0.2162*Temperature + 11.663 $

###d).

The concerns regarding O-rings are justified because there is a statistically significant negative relationship between temperature and O-ring failures. This means that if the temperature is too low, it is very likely that there will be an O-ring failure.


##8.18

###a).

```{r}
p51 = 11.663 - (0.2162 * 51)
p53 = 11.663 - (0.2162 * 53)
p55 = 11.663 - (0.2162 * 55)

probCalc = function(px){
  
  output = exp(px)/(1+ exp(px))
  
  return(output)
  
} 

probCalc(p51)
probCalc(p53)
probCalc(p55)
```

###b).

```{r}
logitcalc = function(x){
  
  px = 11.663 - (0.2162 * x)
  
  output = probCalc(px)
  
  return(output)
  
}

logitx = seq(20,90,1)
logity = sapply(logitx,logitcalc)


plot(logitx,logity, type = "l")
```

###c).

Each predictor is linearly related to logit(p) - This is true and shown by the model output.

Each outcome is independent of all other outcomes - This assumption may be violated since subsequent launches are improved based off of previous launches. It is possible that over time, the technology of the launches are getting better and O-ring failures may be dependent on those improvements.







